---
title: "ABTest"
author: "A Ghimire"
date: "24/12/2024"
output: html_document
---

```

knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.width = 8)
knitr::opts_chunk$set(fig.height = 8)

```

```{r message=FALSE, warning=FALSE}

library(tidyverse)
library(printr)

require(readxl)
require(jpeg)
require(ggimg)

```

# Input and preparation

## Reading Stimulus meta data

```{r}

read_abtest_stim <- function(file = "Stimuli/Stimuli.csv", 
                             stim_dir = "Stimuli/",
                             col_types = cols(File = col_character(),
                                              width = col_double(),
                                              height = col_double(),
                                              mid_x = col_double())) {
  stimuli <- read_csv(file, col_types = col_types) %>% 
    mutate(Path = str_c(stim_dir, File, sep = ""))
  return(stimuli)
}

```  

```{r}

my_stim_col_types <- cols( File = col_character(),
                           width = col_double(),
                           height = col_double(),
                           left_heading = col_character(),
                           left_background = col_character(),
                           right_heading = col_character(),
                           right_background = col_character(),
                           mid_x = col_double())

Stimuli <- read_abtest_stim(file = "Stimuli/Stimuli.csv", col_types = my_stim_col_types)

``` 

## Making a stimulus grid
## Creates a grid of stimuli for visual inspection.

```{r, fig.height = 8, fig.width = 8}

G_0 <- Stimuli %>% 
  ggplot(aes(xmin = 0, xmax = width, ymin = 0, ymax = height)) +
  facet_wrap(~File) +
  ggimg::geom_rect_img(aes(img = Path))

```

## Reading AOI 
##  This function reads Areas of Interest (AOI) data and computes their coordinates relative to each stimulus.

```{r}

read_abtest_aoi <- function(file = "Stimuli/AOI.csv", stim_tab) {
  read_csv(file, 
           col_types = cols(Stimuli = col_character(), AOI = col_character(), 
                            x = col_double(), y = col_double(), 
                            w = col_double(), h = col_double(), Region = col_character())) %>% 
    rename(x_aoi = x, y_aoi = y, w_aoi = w, h_aoi = h) %>% 
    left_join(stim_tab, by = c("Stimuli" = "File")) %>% 
    mutate(xmin = x_aoi, 
           xmax = x_aoi + w_aoi,
           ymax = height - y_aoi, 
           ymin = (height - y_aoi) - h_aoi) %>% 
    arrange(Stimuli, AOI)
}

AOI <- read_abtest_aoi(file = "Stimuli/AOI.csv", stim_tab = Stimuli)

```

## AOI preview
## Visualizes AOIs on top of stimuli images.
```{r}

G_1 <- AOI %>% 
  ggplot(aes(xmin = 0, xmax = width, ymin = 0, ymax = height)) +
  facet_wrap(~Stimuli) +
  ggimg::geom_rect_img(aes(img = Path)) +
  geom_rect(aes(xmin = xmin, ymin = ymin, xmax = xmax, ymax = ymax, fill = AOI),
            alpha = 0.2, inherit.aes = FALSE)

#G_1

```

## Reading Participant Data csv
## Reads and cleans individual participant gaze 

```{r}

read_abtest_csv <- function(file) {
  read_csv(file, 
           col_types = cols(Exp = col_character(), Part = col_character(), 
                            Stimuli = col_character(), time = col_double(), 
                            x = col_double(), y = col_double(), 
                            x_pro = col_double(), y_pro = col_double())) %>% 
    mutate(is_duplicate = x == lag(x) & y == lag(y)) %>% 
    filter(!is_duplicate) %>% 
    mutate(File = file) %>% 
    select(Exp, Part, Stimuli, time, x, y, x_pro, y_pro)
}

```


# Reads and combines multiple participant files, aligning them with Stimuli metadata.

```{r}

read_abtest_files <- function(files, stim_tab) {
  #Data <- 
  files %>% 
    map_df(read_abtest_csv) %>% 
    mutate(Obs = row_number()) %>% 
    mutate(Part = as.factor(as.integer(Part) - min(as.integer(Part)) + 1)) %>% 
    group_by(Part) %>% 
    mutate(time = time - min(time)) %>% 
    ungroup() %>% 
    left_join(stim_tab, by = c("Stimuli" = "Stimuli")) %>% 
    mutate(y = height - y, y_pro = 1 - y_pro) %>% 
    select(Exp, Obs, Part, Stimuli, time, x, y, x_pro, y_pro)
  #return(Data)
}

```

```{r}

csv_files <- dir(path = "Data/", pattern = "*.csv", recursive = TRUE, full.names = TRUE)
Participant_Data <- read_abtest_files(csv_files, Stimuli)

```

## Raw Data Visualization.
## Overlays the last participant's gaze points on the stimuli grid.

```{r}

get_last_part <- function(data) {
  data %>% 
    distinct(Part) %>% 
    filter(as.numeric(Part) == max(as.numeric(Part), na.rm = TRUE)) %>% 
    left_join(data, by = "Part")
  return(last_part)
}


```

```{r}

G_0 +
  geom_point(aes(x = x, y = y),
             size = 2, color = "red", alpha = 0.2,
             data = get_last_part(Participant_Data),
             inherit.aes = FALSE) +
  facet_wrap(~Stimuli)

```

## AOI Classification

```{r}

Participant_Data_AOI <- Participant_Data %>% 
  left_join(AOI, by = "Stimuli") %>% 
  mutate(is_in = x > xmin & x < xmax & y > ymin & y < ymax) %>% 
  filter(is_in) %>% 
  select(Obs, AOI) %>% 
  right_join(Participant_Data, by = "Obs") %>% 
  mutate(AOI = if_else(is.na(AOI), "Outside", AOI)) %>% 
  arrange(Part, time)

summary(Participant_Data_AOI)

```

## Measuring Visits

## Derive Visits Based on Travel Threshold

```{r}

travel_threshold <- 50

Participant_Visits <- Participant_Data_AOI %>% 
  group_by(Part, Stimuli) %>% 
  filter(AOI != lag(AOI) | travel > travel_threshold) %>% 
  mutate(visit = row_number(),
         duration = lead(time) - time) %>% 
  ungroup()

```

## Participant-level Summary

### Frequencies and durations

```{r}

Participant_Summary <- Participant_Visits %>% 
  group_by(Part, AOI, Region) %>% 
  summarize(n_visits = n(),
            total_dur = sum(duration, na.rm = TRUE)) %>% 
  ungroup

```

## Aggregate Fixation Durations by Color

```{r}

color_attention <- Participant_Data_AOI %>% 
  group_by(AOI) %>% 
  summarize(total_dur = sum(duration, na.rm = TRUE)) %>%  # Summing fixation durations for each AOI
  ungroup() %>% 
  separate(AOI, into = c("Color", "Region"), sep = "_") %>%  # Split AOI into color and region
  group_by(Color) %>% 
  summarize(color_total_dur = sum(total_dur))  # Aggregate across regions

```

# Visualize the total fixation duration by color

```{r}

color_attention %>% 
  ggplot(aes(x = Color, y = color_total_dur, fill = Color)) +
  geom_col() +
  labs(title = "Total Fixation Duration by Color", 
       x = "Color", 
       y = "Total Fixation Duration") +
  theme_minimal()

```

## Examine Fixation Trends by Color and Region

```{r}

color_region_attention <- Participant_Data_AOI %>% 
  group_by(AOI) %>% 
  summarize(total_dur = sum(duration, na.rm = TRUE)) %>% 
  ungroup() %>% 
  separate(AOI, into = c("Color", "Region"), sep = "_") %>% 
  group_by(Color, Region) %>% 
  summarize(region_total_dur = sum(total_dur))  # Aggregate fixation durations by color and region

```

# Visualize fixation duration by color and region

```{r}

color_region_attention %>% 
  ggplot(aes(x = Region, y = region_total_dur, fill = Color)) +
  geom_col(position = "dodge") +
  labs(title = "Fixation Duration by Color and Region", 
       x = "Region", 
       y = "Total Fixation Duration") +
  theme_minimal()
```

## Visualization of Fixation Durations by AOI

```{r}

G_2 <- Participant_Summary %>% 
  ggplot(aes(x = AOI, y = total_dur, fill = Region)) +
  facet_wrap(~Part) +
  geom_col()

G_2

```


```{r}

save(AOI, Participant_Data, Participant_Data_AOI, color_attention, color_region_attention, Participant_Visits, Participant_Summary, file = "abtest_results.Rda")

```